---
title: "Analysis of NOAA Storm Database and its impact on human life."
output: html_document
---

## Synopsis

## Data Processing

Start by downloading and extracting the csv file from the bz2 archive file. The
data is stored as a data frame `storm.data`.

```{r cache=TRUE}
zipFileName <- "repdata-data-StormData.csv.bz2"
con <- bzfile(zipFileName)
storm.data <- read.csv(con)
```

We can get a summary of all the different event types by using the `summary`
function on the column `EVTYPE

```{r}
summary(storm.data$EVTYPE)
```

Using `head` we can see the top 10 weather events in the data frame,

```{r}
head(summary(storm.data$EVTYPE), n=10)
```

From this we can see that due to some inconsistance in the labelling we have
some data that is similar but not the same but could be counted as the same. For
example `THUNDERSTORM WIND` and `THUNDERSTORM WINDS` are clearly the same. Is
this the same as `HIGH WIND`. It is almost certainly the same as `TSTM WIND`.

So lets clean up these `EVTYPE`s.

```{r}
storm.data$EVTYPE <- gsub("TSTM","THUNDERSTORM",storm.data$EVTYPE)
storm.data$EVTYPE <- gsub("WINDS","WIND",storm.data$EVTYPE)
storm.data$EVTYPE <- gsub("MARINE ","",storm.data$EVTYPE)
storm.data$EVTYPE <- gsub("FLASH ","",storm.data$EVTYPE)
storm.data$EVTYPE <- gsub("STREAM FLD", "FLOOD", storm.data$EVTYPE)
storm.data$EVTYPE <- gsub("URBAN/SML FLOOD", "FLOOD", storm.data$EVTYPE)
storm.data$EVTYPE <- gsub("FLOODING", "FLOOD", storm.data$EVTYPE)
storm.data$EVTYPE <- gsub("COASTAL FLOOD", "FLOOD", storm.data$EVTYPE)
storm.data$EVTYPE <- gsub("STRONG WIND", "HIGH WIND", storm.data$EVTYPE)
storm.data$EVTYPE <- gsub("WILD/FOREST FIRE","WILDFIRE",storm.data$EVTYPE)
storm.data$EVTYPE <- factor(storm.data$EVTYPE)
head(summary(storm.data$EVTYPE),n=15)
```

We now need to count the number of fatalities and injuries by the `EVTYPE`.

## Results